---
title: "Bellabeat"
author: "Beau"
date: "`r Sys.Date()`"
output: html_document
---
## The Bellabeat Case Study
For this case study, I'm being asked by the wearable tech company Bellabeat to analyze smart device usage data in order to gain insight into how consumers use non-Bellabeat smart devices. They then want me to select one Bellabeat product to apply these insights to in a presentation. These questions will guide my analysis: 

1. What are some trends in smart device usage?
2. How could these trends apply to Bellabeat customers?
3. How could these trends help influence Bellabeat marketing strategy?
This report will include the following deliverables:
1. A clear summary of the business task
2. A description of all data sources used
3. Documentation of any cleaning or manipulation of data
4. A summary of my analysis
5. Supporting visualizations and key findings
6. My top high-level content recommendations based on my analysis

## Stating the buisiness task
The business task is to identify smart device usage trends and theorize how hose trends could apply to current customers and influence our marketing strategy in order to gain new customers. 

## Description of all data sources used
The data I have been given to be used is the [FitBit Fitness Tracker Data](https://www.kaggle.com/arashnic/fitbit). This is a publicly available dataset in which 30 eligible Fitbit users consented to the submission of personal tracker data. The data includes daily steps, sleep, heart rate, etc. 
Privacy is insured through the use of an anonymous user ID number instead of a name so there is very low risk of sharing personally identifiable information. This data is six years old which is not current. There may be biased data due to the fact that age and gender are not reported in the dataset. The data in some categories may be useless due to the fact that some categories (such as calories) are self-reported by each user. The license for this data is public, with no known security issues. This dataset has a low number of users, which may skew the data.

## How to determine exercise
Seeing as different individuals achieve fitness goals in different ways, I tried to find a more universal measure for public health. In my research, I found a metric included in the dataset which every user reported doing: daily steps. I found this article by the US government's [National Institute of Health](https://www.nih.gov/news-events/nih-research-matters/number-steps-day-more-important-step-intensity) which supported my idea with data less than three years old. The evidence was also supported by [this article](https://jamanetwork.com/journals/jamainternalmedicine/fullarticle/2734709?guestAccessKey=4d4da1ad-4728-48cf-b424-1cd0523b6f0d&utm_source=For_The_Media&utm_medium=referral&utm_campaign=ftm_links&utm_content=tfl&utm_term=052919) from the Journal of the American Medical Association.

## Documentation of cleaning/manipulation of data
Due to ease of use, I began the cleaning process in Google Sheets. I downloaded the dataset. Then I changed the time to a 24 hour format using the DateTime function. Next, I separated date and time into their own columns and named those columns 'Date' and 'Time'. I then used the remove duplicates function to deduplicate the data. Following that, I changed each row which contained decimal places to a format which only had two decimal places to increase readability. During my cleaning, I noticed that the weight log only had eight unique users and sleep day only had 25 unique users while each of the other datasets had 33 unique users. Also, I noticed a lot of zero values in the dataset, indicating that users were not reporting the data or not using the fitness tracker. After cleaning the data, I decided to finish my analysis in RStudio so that I could produce customized visualizations as well as produce a report by using a .rmd file.

## Setting up the R environment
Notes: setting up my R environment by loading 'tidyverse', 'googlesheets4', 
'googledrive', and 'dplyr' packages

```{r loading packages, message=FALSE, warning=FALSE}
library(googlesheets4)
library(tidyverse)
library(dplyr)
library(googledrive)

```
## Reading the data from Google Sheets
```{r message=FALSE, warning=FALSE}

daily_activities<-read_sheet("https://docs.google.com/spreadsheets/d/1gYMmHeNXB0RfNTvoM-Cr7wmvKHwuce0AMdur7sBPsAI/edit#gid=787649191")

```

## Visualizations
Here we will go through our visualizations

## Scatterplot of daily steps over time
Here, we plot how many steps each user recorded over time with a regression line. Please note the regression line remains virtually straight, only trending downward at the end of the time period. This downward trend may indicate a lack of exercise or a lack of using the Fitbit properly.
```{r message=FALSE, warning=FALSE}
ggplot(data=daily_activities, aes(x=Date, y=TotalSteps)) + 
  geom_point(color = "Red") +
  geom_smooth()
```

## Scatterplot of sedentary minutes over time
Here, we plot how many minutes per day each user recorded themselves as sedentary. Please note the regression line curves sharply downward towards the end of the time period, with one user even recording zero sedentary minutes. This further indicates that something is wrong with the data collection.
```{r message=FALSE, warning=FALSE}
ggplot(data=daily_activities, aes(x=Date, y=SedentaryMinutes)) + 
  geom_point(color = "Red") +
  geom_smooth()
```

## Scatterplot of calories over time
Here, we plot how many calories each user recorded consuming over time. Please note the downward curve in the regression line at the end of the time period, as well as the increase in users reporting consuming fewer than 1000 calories per day, indicating that they are starving themselves or they are not reporting their calories accurately.
```{r message=FALSE, warning=FALSE}
ggplot(data=daily_activities, aes(x=Date, y=Calories)) + 
  geom_point(color = "Red") +
  geom_smooth()
```

## Conclusions
Due to the decrease in daily steps, sedentary minutes, and calories, I have come to the conclusion that the usage of this wearable tech declined over time. Perhaps people got bored or just decided to give up recording their activity. One would expect that the daily steps and sedentary minutes to trend inversely to one another, but in this case they did not. Also, the increase in users reporting consuming fewer than 1000 calories is concerning. 
Other explanations include fasting combined with an increase in activities which do not require any walking. 
The Bellabeat marketing team may want to invest in selling more Bellabeat memberships to ensure that users continue to record their progress while on their fitness journey. This additional contact will not only provide better data to make business decisions, but could provide a long-term source of profits due to high customer engagement. 